---
title: Are there clusters?
subtitle: Ziv Edits Version 0.3.1
author:
- <h5 style="font-style:italic"> Ziv Ben Zion
- <h5 style="font-style:italic"> Tobias R. Spiller
output: html_document
---

This uses the methods shown in Dinga et al. (2019) [10.1016/j.nicl.2019.101796. The Original code can be found: https://github.com/dinga92/niclin2019-biotypes/blob/master/analysis_code.Rmd.
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Libraries , include=FALSE}
library(readxl)
library(PerformanceAnalytics)
library(corrplot)
library(NbClust)
library(MASS)
library(Hmisc)
library(cluster)
library(factoextra)
```

# Import data
```{r Import Neural Data, include=FALSE}
roi_data <- read_excel("roi_data_ziv.xlsx")#158 subjects at TP1 with at least one task
roi_data <- na.omit(roi_data) #remove 28 cases with any missing data (n=3 without Hariri & n=25 without Domino) 
roi_data_raw <- roi_data 
```

# Prepare data
## Remove Outliers
```{r Outliers Correction}
chart.Correlation(roi_data_raw[,2:8], histogram=TRUE, pch=19)
roi_datafz1<- roi_data_raw
roi_datafz1[,c(2:8)] <- scale(roi_datafz1[,c(2:8)]) #scale to z-distribution
roi_datafz1[,2:8][roi_datafz1[2:8]>=3] <- 3 #replace outliers by capping to +3
roi_datafz1[,2:8][roi_datafz1[2:8]<=-3] <- -3 #replace outliers by capping -3
```

# Analysis
## Stage #1 - Covariance among the fMRI tasks and ROIs
```{r Correlation Matrices}
data_cor <- roi_datafz1[,2:8]
# Rename
colnames(data_cor) <- c("T_Insula", "T_dACC", "T_Amy", "T_sgACC",
                        "R_Amy", "R_NAcc", "R_OFC")
# Correlaion matrix
mat1 <- Hmisc::rcorr(as.matrix(data_cor))
mat1
# Save P-Value
mat1p <- mat1$P
# Correlaion matrix only complete data
mat2 <- cor(data_cor, use="complete.obs")
round(mat2, 2)
```

## Stage #2 - Hierarchical clustering analysis
```{r }
#performing clustering
hc1 <- agnes(roi_datafz1[2:8], metric = "euclidean", method = "ward")
row.names(hc1$data) <-c(1:130)

#show the agglomerative coefficient, which measures the amount of clustering structure found (values closer to 1 suggest strong clustering structure).
hc1$ac 

#plot tree of hierarchical clustering
#pdf(file = "C:/Users/zhb4/Box Sync/Brain based Biotypes (Jennifer Stevens)/Clustering_Tree.pdf", width = 4, height = 4) 
pltree(hc1, cex = 0.6, hang = -1, main = "Clustering")
#dev.off()
```

### Determine optimal number of clusters using different metrics
```{r Optimal Number of Clusters}
#Hartigan
nb1 <- NbClust(roi_datafz1[,2:8], distance = "euclidean", min.nc = 2, max.nc = 6, method = "ward.D2", index="hartigan") # Estimate
fviz_nbclust(nb1) # Visualize
# Silhouette
nb2 <- NbClust(roi_datafz1[,2:8], distance = "euclidean", min.nc = 2, max.nc = 6, method = "ward.D2", index="silhouette") # Estimate
fviz_nbclust(nb2) # Visualize
```

## Stage #3 - Examining the statistical significance of the clusters
### Range of clusters 2 to 6

3a. Make a function that performs a hierarchical clustering and return the highest clustering indexes
```{r, message=F}
#create a new dataset excluding the subject numbers (column #1) 
test_data <- roi_datafz1[,2:8]
#Make a new function that performs a hierarchical clustering and return the highest clustering indexes
cluster_test <- function(test_data){
  hcfit_hart <- NbClust(test_data, method="ward.D2", index="hartigan", min.nc=2, max.nc = 6)
  Hartigan_index <- max(hcfit_hart$All.index)
  hcfit_Sil <- NbClust(test_data, method="ward.D", index="silhouette", min.nc=2, max.nc = 6)
  Silhouette_index <- max(hcfit_Sil$All.index) 
  return(c("Hartigan"= Hartigan_index , "Silhouette"= Silhouette_index))
}
```

3b. Simulate multivariate normally distributed data from the covariance matrix extracted from the original data used to perform hierarchical clustering
```{r}
sigma <- cov(test_data)
mu <- colMeans(test_data)
real_CI <- cluster_test(test_data) #Hartigan and Silhouette max values for the real clusters
```

3c. Repeatedly perform hierarchical clustering on samples from this distribution, thus creating an empirical null distribution of clustering indeces
```{r}
# get a null distribution of clusters
null_Cluster <- list()
n_sims <- 1999 #number of simulations (1,999)
for (i in 1:n_sims){
  rand_sample <- mvrnorm(n=nrow(test_data), mu=mu, Sigma=sigma) #multivariate normally distributed data
  null_Cluster[[i]] <- cluster_test(rand_sample)
}
null_Cluster <- as.data.frame(do.call(rbind, null_Cluster))#transform results to data frame
```

3d. print p-values
```{r}
#Hartigan
rank_cv1 <- sum(real_CI[1] < null_Cluster[,1]) + 1 # total of 2000
pval_cv1 <- rank_cv1 / (n_sims+1)
#Silhouette
rank_cv2 <- sum(real_CI[2] < null_Cluster[,2]) + 1 # total of 2000
pval_cv2 <- rank_cv2 / (n_sims+1)
t(t((c("p.val variance ratio"=pval_cv1, "p.val Silhouette"=pval_cv2))))#transpose the p-values
```

3e. visualize null distribution
```{r}
par(mfrow=c(1,2))
hist(null_Cluster[,1], breaks = 30, main = "Hartigan distance null")
abline(v=real_CI[1], col="red")
text(real_CI[1] + 10, 70, paste('p = ', round(pval_cv1, 2)))
hist(null_Cluster[,2], breaks = 30, main = "Silhouette null")
abline(v=real_CI[2], col="red")
text(real_CI[2] - 0.025, 80, paste('p = ', round(pval_cv2, 2)))
```
