---
title: "Test"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Libraries , include=FALSE}
library(readxl)
library(PerformanceAnalytics)
library(Hmisc)
library(corrplot)
library(factoextra)
library(cluster)
library(factoextra)
library(NbClust)
library(dendextend)
library(ggplot2)
library(writexl)
library(readxl)
library(ggpubr)
library(reshape2)
library(plyr)
library(fpc) 
library(dendextend)
library(class)
library(gmodels)
library(caret)
library(lme4)
library(lmerTest)
library(sjstats)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(dplyr)
#library(ggradar)
#library(ggboxplot)
```

```{r Import Neural Data, include=FALSE}
roi_data <- read_excel("roi_data_ziv.xlsx")#158 subjects at TP1 with at least one task
roi_data <- na.omit(roi_data) #remove 28 cases with any missing data (n=3 without Hariri & n=25 without Domino) 
roi_data_raw <- roi_data 
```

```{r Outliers Correction}
chart.Correlation(roi_data_raw[,2:8], histogram=TRUE, pch=19)
roi_datafz1<- roi_data_raw
roi_datafz1[,c(2:8)] <- scale(roi_datafz1[,c(2:8)]) #scale to z-distribution
roi_datafz1[,2:8][roi_datafz1[2:8]>=3] <- 3 #replace outliers by capping to +3
roi_datafz1[,2:8][roi_datafz1[2:8]<=-3] <- -3 #replace outliers by capping -3
```

Stage #1 - Covariance among the fMRI tasks and ROIs
```{r Correlation Matrices}
data_cor <- roi_datafz1[,2:8]
colnames(data_cor) <- c("T_Insula", "T_dACC", "T_Amy", "T_sgACC",
                                                  "R_Amy", "R_NAcc", "R_OFC")

mat1 <- rcorr(as.matrix(data_cor))
mat1
mat1p <- mat1$P
mat2 <- cor(data_cor, use="complete.obs")
round(mat2, 2)
```
Stage #2 - Clustering analysis based on task-based fMRI indices
```{r }
#performing clustering
hc1 <- agnes(roi_datafz1[2:8], method = "ward")
row.names(hc1$data)<-c(1:130)

#show the agglomerative coefficient, which measures the amount of clustering structure found (values closer to 1 suggest strong clustering structure).
hc1$ac 

#plot tree of hierarchical clustering
#pdf(file = "C:/Users/zhb4/Box Sync/Brain based Biotypes (Jennifer Stevens)/Clustering_Tree.pdf", width = 4, height = 4) 
pltree(hc1, cex = 0.6, hang = -1, main = "Clustering")
#dev.off()

```

```{r Optimal Number of Clusters Hartigan’s distance index}
nb1 <- NbClust(roi_datafz1[,2:8], distance = "euclidean", min.nc = 2,max.nc = 6, method = "ward.D2", index="alllong")
fviz_nbclust(nb1)

#pdf(file = "C:/Users/zhb4/Box Sync/14. Brain based Biotypes (Jennifer Stevens)/Output/Supplementary_Figure_1a.pdf", width = 4, height = 4) 
fviz_nbclust(roi_datafz1[,2:8], hcut, method = "wss", k.max=6) + geom_vline(xintercept = 4, linetype = 2, color="steelblue")
#dev.off()
```

```{r Optimal Number of Clusters Silhouette}
#pdf(file = "C:/Users/zhb4/Box Sync/14. Brain based Biotypes (Jennifer Stevens)/Output/Supplementary_Figure_1b.pdf", width = 4, height = 4) 
fviz_nbclust(roi_datafz1[,2:8], hcut, method = "silhouette", k.max=6)
#dev.off()
```

```{r Optimal Number of Clusters GAP Statistics}
gap_stat <- clusGap(roi_datafz1[,2:8], FUN = hcut, K.max = 6, nboot=500)
fviz_gap_stat(gap_stat)
```

Stage #3 - Examining the statistical significance of the clusters


3a. Make a function that performs a hierarchical clustering and return the highest clustering indexes
```{r, message=F}
#create a new dataset excluding the subject numbers (column #1) 
test_data <- roi_datafz1[,2:8]

#Make a new function that performs a hierarchical clustering and return the highest clustering indexes
#My changes - Hartigan instead of CH, ward.D2 instead of ward.D, min 2 instead of 3, max 6 instead of 5
cluster_test <- function(test_data){
  hcfit <- NbClust(test_data, method="ward.D2", index="hartigan", min.nc=2, max.nc = 6)
  Hartigan_index <- max(hcfit$All.index)
  hcfit <- NbClust(test_data, method="ward.D", index="silhouette", min.nc=2, max.nc = 6)
  Silhouette_index <- max(hcfit$All.index) 
  return(c("Hartigan"=Hartigan_index , "Silhouette"= Silhouette_index))
}

#Optimal Number of Clusters Hartigan’s distance index (from my original script)
#nb1 <- NbClust(roi_datafz1[,2:8], distance = "euclidean", min.nc = 2,max.nc = 6, method = "ward.D2", index="alllong")
#fviz_nbclust(nb1)
#nb1$All.index[3,3]#Hartigan for 4 clusters
#nb1$All.index[3,13]#Silhouette for 2 clusters
```

3b. Fit a multivariate normal distribution to the same data used to perform hierarchical clustering
```{r}
library(MASS)
sigma <- cov(test_data)
mu <- colMeans(test_data)
real_CI <- cluster_test(test_data)#Hartigan and Silhouette max values for the real clusters
```

3c. Repeatedly perform hierarchical clustering on samples from this distribution, thus creating an empirical null distribution of clustering indeces
```{r}
# get a null distribution of clusters
null_Cluster <- list()
n_sims <- 1999 #number of simulations
for (i in 1:n_sims){
  rand_sample <- mvrnorm(n=nrow(test_data), mu=mu, Sigma=sigma)#multivariate normally distributed data
  null_Cluster[[i]] <- cluster_test(rand_sample)
}
null_Cluster <- as.data.frame(do.call(rbind, null_Cluster))#transform results to data frame
```

3d. print p-values

```{r}
#Hartigan
rank_cv1 <- sum(real_CI[1] < null_Cluster[,1]) + 1 #in the begining it is 6, after changing to 1999 sim it will be 2000 total
pval_cv1 <- rank_cv1 / (n_sims+1)
#Silhouette
rank_cv2 <- sum(real_CI[2] < null_Cluster[,2]) + 1
pval_cv2 <- rank_cv2 / (n_sims+1)

t(t((c("p.val variance ratio"=pval_cv1, "p.val Silhouette"=pval_cv2))))#transpose the p-values
```

3e. visualize null distribution

```{r}
par(mfrow=c(1,2))
hist(null_Cluster[,1], breaks = 30, main = "variance ratio criterion null")
abline(v=real_CI[1], col="red")
text(real_CI[1] + 10, 70, paste('p = ', round(pval_cv1, 2)))
hist(null_Cluster[,2], breaks = 30, main = "Silhouette null")
abline(v=real_CI[2], col="red")
text(real_CI[2] - 0.025, 80, paste('p = ', round(pval_cv2, 2)))
```


